An Analysis of Health & Lifestyle Factors Associated with Sleep Quality

### 2024-05-02 products/manuscript/Manuscript.qmd
pushing for zane to look

---
title: "An Analysis of Health & Lifestyle Factors Associated with Sleep Quality"
subtitle: ""
author: Kevin Kosewick and Emma Hardin-Parker
date: "4-29-2024"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: sleepref2.bib
csl: ../../assets/apa.csl
---

@@

@@

@@

![Figure 6: The ranking of each variable's importance in determining
sleep quality according to our Random Forest
model.](images/variableimportance.png)

As seen in figure 6, Sleep Duration, Stress Level, and Age had the
highest levels of importance in our model. Physical Activity Level was
excluded entirely, which was to be expected given our exploratory data
analysis and simple linear regression models.

@@

The differences between the null model and RF model's metrics show that
our RF model is better suited for predicting sleep quality on unseen
data than a model using no predictors. The R-squared value of 0.997 is
extremely high; this shows that a large amount of variation is explained
using our RF model. The variable importance plot we created answers our
questions about which variables are most important for predicting sleep
quality. As we can see, sleep duration, stress level, and age were the
top three predictors for sleep quality in our data. This reflects the
relationships we saw in our colinearity plot as well as the clear
correlations we found between the variables in our EDA. We did not
expect to see age so high on the list; this could be due to us grouping
ages together in our EDA and therefore misleading us in our preliminary
data exploration. We did not use the groups when creating our models as
we learned that grouping continuous variables is a poor practice as we
progressed in our MADA course.

@@

>>>>

---
title: "An Analysis of Health & Lifestyle Factors Associated with Sleep Quality"
subtitle: ""
author: Kevin Kosewick and Emma Hardin-Parker
date: "4-29-2024"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/sleepref2.bib
csl: ../../assets/apa.csl
---

@@

For our LASSO Regression model, used a 5-fold Cross-Validation repeated 5 times 
to train our data. We used a tuning parameters and penalties within the tuning
grid. Using a regularization method for this analysis was something we were 
concerned about due to the high level of correlation between predictors;
however, we wanted to include it to ensure a more robust analysis of our data.

@@

![Figure 6: LASSO Tuning Results](images/lasso_tuneresults.png)

Figure 6 suggests that regularization may not be the best fit for our data. 
Though the RMSE values are small, we should be careful in analyzing this model
due to the high level of correlation and collinearity between variables.

![Figure 7: The observed vs. predicted values for our 
"LASSO Model.](images/lasso_obsvpred.png)

We can see in Figure 7 that the LASSO Model did decently well predicting on 
the new data. The LASSO model can sufficiently predict sleep quality using
unseen data.

@@

![Figure 8: The ranking of each variable's importance in determining
sleep quality according to our Random Forest
model.](images/variableimportance.png)

As seen in Figure 8, Sleep Duration, Stress Level, and Age had the
highest levels of importance in our model. Physical Activity Level was
excluded entirely, which was to be expected given our exploratory data
analysis and simple linear regression models.

@@

The differences between the null model and RF model's metrics show that
our RF model is better suited for predicting sleep quality on unseen
data than a model using no predictors. The R-squared value of 0.997 is
extremely high; this shows that a large amount of variation is explained
using our RF model. The variable importance plot we created answers our
questions about which variables are most important for predicting sleep
quality. As we can see, sleep duration, stress level, and age were the
top three predictors for sleep quality in our data. This reflects the
relationships we saw in our colinearity plot as well as the clear
correlations we found between the variables in our EDA. We did not
expect to see age so high on the list; this could be due to us grouping
ages together in our EDA and therefore misleading us in our preliminary
data exploration. We did not use the groups when creating our models as
we learned that grouping continuous variables is a poor practice as we
progressed in our MADA course. Our RF model performed better than the LASSO 
model, which was to be expected. We struggled to create a variable importance
plot for the LASSO model, which could be due to sensitive parameters and 
tuning penalties.

@@

Additionally, while we conducted a LASSO regression model and received favorable 
results, regularization methods are difficult to conduct and interpret when 
predictors are highly correlated. The penalty paramaters appeared to be too 
sensitive and were affected by the collinearity between predictors.
